---
title: "Final_Project"
author: "Rom Amsili, Oryan Yehezkel, Din Ezra"
date: "316509397, 311495824, 206065989"
output: html_document
---

```{r setup_Libraries, include=FALSE, echo=FALSE}
library(tidyverse)
library(dplyr)
library(usmap)
library(zoo)
library(reshape2)
library(rvest)
library(ggpubr)
library( taRifx )
library(SciViews)
library(gganimate)
library(rgdal)
library(lubridate)
library(hrbrthemes)
library(janitor)
library(kableExtra)
```

```{r load_CSVs, include=FALSE, echo=FALSE}
# load CSVs

concern <- read.csv("civiqs_poll.csv")
confirm <- read.csv("table-1.csv")
death <- read.csv("table-2.csv")
# page<- read_html("https://en.wikipedia.org/wiki/Template:COVID-19_pandemic_data/United_States_medical_cases")
# confirm <- page %>% html_node(xpath = '//*[@id="mw-content-text"]/div[1]/div[2]/table') %>%
#   html_table(fill = TRUE)
# 
# death<- page %>%
#   html_node(xpath = '//*[@id="mw-content-text"]/div[1]/div[2]/table')%>%
#   html_table(fill = TRUE)
```

We have got 3 Rows(Raw) Tables:\
concern:\
```{r concern_table, echo=FALSE}
colnames(concern)[1] <- "Date"
head(concern)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
confirm:\
```{r confirm_table, echo=FALSE}
head(confirm[c(1:6,57:59)])%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
death:\
```{r death_table, echo=FALSE}
head(death[0:10])%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
we creted new table: loc\
```{r,creating a new table Country-Region, echo=FALSE}
# creating a new table Country-Region

loc <- data.frame(t(colnames(confirm)))
names(loc) <- t(confirm[1, ])
loc <- loc[,-1]
loc<- loc[,0:56]
loc <- loc%>%select(-Date)%>%pivot_longer(
  cols = AK:VI,
  names_to = "Country",
  values_to = "Region"
)
loc <- loc%>%mutate(Region = sub('\\.\\d*',"",Region))
head(loc)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
## Cleaning Table1 - confirms:
* the first row will be the header\
* replace NA with 0\
* filtered the rows with only dates and removing unnecessary rows\
* apply date\numerical type on variables\
```{r set Columns Names and NA, include=FALSE}
# set Columns Names

colnames(confirm) = confirm[1, ] # the first row will be the header
confirm <- confirm %>% replace(.,.=="","0") # replace empty string with "0"

```
split into 2 tables: per_country and daily_total\
```{r split into 2 tables: per_country and daily_total, include=FALSE}
# split into 2 tables: per_country and daily_total

confirm_total<- confirm[,57:64]
confirm_dates<- confirm[,0:56]
```

```{r filtered the rows with only dates and removing unnecessary rows, include=FALSE}
# filtered the rows with only dates and removing unnecessary rows

colnames(confirm_total) = c("Date","Daily_confirmed","Total_confirmed","Daily_deaths","Total_deaths","Daily_recovered","Total_recovered","Active")
confirm_dates <- confirm_dates %>% filter(Date %in% c(Date %>% str_subset("-")))
confirm_total <- confirm_total %>% filter(Date %in% c(Date %>% str_subset("-")))

```

```{r apply date\numerical type on variables, include=FALSE}
# date or numerical type on variables and change to positive numbers only

confirm_dates<- confirm_dates %>% mutate(Date = as.Date(Date, format = "%d-%b-%y"))
confirm_total<- confirm_total %>% mutate(Date = as.Date(Date, format = "%d-%b-%y"))
confirm_total[,2:8] <- lapply(confirm_total[,2:8],function(x){abs(as.numeric(gsub(",", "", x)))})
confirm_dates[,2:56] <- lapply(confirm_dates[,2:56],function(x){abs(as.numeric(gsub(",", "", x)))})
```

```{r, echo=FALSE}
head(confirm_dates[0:10])%>% kbl() %>%
  kable_material(c("striped", "hover"))
head(confirm_total)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
# create Table confirm_dates_longer, its will help us with aggregation and group_by functions.
```{r create confirm_dates_longer, echo=FALSE}
# create Table confirm_dates_longer, its will help us with aggregation and group_by functions.

confirm_dates_longer <- confirm_dates%>%pivot_longer(
  cols = AK:VI,
  names_to = "Country",
  values_to = "confirms"
)
confirm_dates_longer <- confirm_dates_longer%>%mutate(
  Country = as.factor(Country),
  confirms = as.numeric(confirms))
head(confirm_dates_longer)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
adding Region to confirm_dates_longer table.\
```{r adding Region to confirm_dates_longer table, message=FALSE, echo = FALSE}
# adding Region to confirm_dates_longer table

confirm_dates_longer <- left_join(confirm_dates_longer,loc)
confirm_dates_longer <- confirm_dates_longer %>%relocate(Region,.before = Country)
head(confirm_dates_longer)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
## Exploring Table2 - Deaths:
same way as Confirms table _ Death_Longer.\
```{r,cleaning 2nd table, echo=FALSE}
# cleaning 2nd table
# write.csv(confirm_total,"death_total.csv")
colnames(death) = death[1, ]
death <- death %>% replace(.,.=="","0")
death <- death[,0:56]
death <- death %>% filter(Date %in% c(Date %>% str_subset("-")))
death <- death %>% mutate(Date = as.Date(Date, format = "%d-%b-%y"))
```

```{r,change to numeric type, include=FALSE}
# change to numeric type

death[,2:56] <- lapply(death[,2:56],function(x){abs(as.numeric(gsub(",", "", x)))})

```

```{r create death_longer, message=FALSE,echo=FALSE}
# create death_longer

death_longer <- death%>%pivot_longer(
  cols = AK:VI,
  names_to = "Country",
  values_to = "Death"
)
death_longer <- death_longer%>%mutate(
  Country = as.factor(Country),
  Death = abs(as.numeric(Death)))
death_longer <- left_join(death_longer,loc)
death_longer <- death_longer %>%relocate(Region,.before = Country)
head(death_longer)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
# Join BTW death&confirm
```{r join BTW death&confirm, message=FALSE, echo=FALSE}
# join BTW death&confirm

death_confirm <- full_join(death_longer, confirm_dates_longer) 
death_confirm[is.na(death_confirm)] <- 0
head(death_confirm)%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
## Features of the data:
Basic information:\
```{r Features of the data, echo = FALSE}
# Minimum and Maximum Deaths and Confirms

mean_general  <-    death_confirm %>% 
                      summarise(
                        Type = c("Death","Confirms"),
                        min = c(min(Death),min(confirms)),
                        max = c(max(Death),max(confirms)),
                        mean = trunc(c(mean(Death),mean(confirms))),
                        median = trunc(c(median(Death),median(confirms))))
mean_general <- mean_general %>% mutate(std = trunc(c(sd(death_confirm$Death),sd(death_confirm$confirms))))
mean_general%>% kbl() %>%
  kable_material(c("striped", "hover"))

mean_by_country_deaths <- aggregate(death_confirm[, 4:5], list(death_confirm$Country), mean)
mean_by_country_deaths <-  mean_by_country_deaths[order(c(mean_by_country_deaths$confirms,mean_by_country_deaths$death),decreasing = TRUE),]
mean_by_country_deaths <- mean_by_country_deaths %>% mutate(Death = trunc(Death),
                                  confirms = trunc(confirms))
```
showing the 5 most countries with the highest daily result.\
```{r, echo = FALSE}
# showing the 5 most countries with the highest daily result.
five_countries_most_deaths  <- head(mean_by_country_deaths,5)
colnames(five_countries_most_deaths)[1] <- "State"
five_countries_most_deaths[1:3]%>% kbl() %>%
  kable_material(c("striped", "hover"))
```
## Visualization\
```{r Visualization1, message=FALSE, warning=FALSE, echo = FALSE}
death_confirm_5 <- death_confirm %>%
                      filter(Country %in% c(five_countries_most_deaths$State))

ggplot(death_confirm_5, aes(y= confirms, x=Date, color = Country))+
              geom_smooth(aes(color = Country, group = Country),span = 0.25)+
              geom_hline(yintercept=0, linetype="dashed", color = "red")+
              labs(title = "Daily Confirms in the 5 most infected countries", x = "Date", y = "Daily Confirms",caption = "We can see the Waves of the Epidemic") + 
              facet_wrap(~Country, ncol=2)

```

## Location\
\
\
```{r, echo = FALSE}
temp_countries <- death_confirm_5 %>% group_by(Country) %>% summarise(n=1)
colnames(temp_countries)[1] <- "state"
plot_usmap(regions = "state", data = temp_countries, values = "n") +
  theme(panel.background = element_rect(colour = "black"))+
  labs(title = "5 Most infected Countries Location", caption = "spread all over USA")+
    theme(legend.position = "none")
```
\
## Compare Between Confirms & Deaths\
\
\
```{r Visualization, message=FALSE, warning=FALSE,echo = FALSE}
death_confirm_CA_most <- death_confirm_5 %>% filter(Country == "CA")
coef <- 764/53711
ggplot(death_confirm_CA_most, aes(x=Date))+
              geom_smooth(aes(y= confirms),span = 0.25)+
              geom_smooth(aes(y= Death/coef),span = 0.25,linetype = "dashed",color = "steelblue1")+
              geom_hline(yintercept=0, linetype="dashed")+
  scale_y_continuous(
    
    # Features of the first axis
    name = "Confirms",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coef, name="Deaths")
  ) + 
  
  theme_ipsum() +

  theme(
    axis.title.y = element_text(color = "black", size=13),
    axis.title.y.right = element_text(color = "steelblue1", size=13),
    axis.line.y.right = element_line(color = "steelblue1",linetype="dashed",size = 1),
    axis.text.y.right = element_text(color = "steelblue1"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  labs(title = "California Daily Confirms & Deaths",
       x = "Date",caption = "the pick of Deaths come after the pick of confirms")
```
\
\
```{r, message=FALSE, warning=FALSE, echo = FALSE}
regions <- death_confirm %>% 
  mutate(month = month(Date),
         year = year(Date)) %>% 
  group_by(Region, year,  month) %>% 
  summarise(Deaths = sum(Death),Confirms = sum(confirms)) %>%
  mutate(Date = as.Date(paste(year, month, 1,sep="-"), "%Y-%m-%d"))
  
# regions <- regions%>%pivot_longer(
#   cols = Deaths:Confirms,
#   names_to = "D_C",
#   values_to = "count"
# )
ggplot(data = regions,aes(x = Date,y = Deaths, color = Region))+
  geom_smooth(position="dodge",span = 0.25)+
  theme(legend.position = "bottom")+
  labs(title = "Monthly Deaths in the Regions", x = "Date", y = "Deaths")

```
\
the majority: the second wave was with more deaths cases then the first one.\
\
\

```{r visualization_Oryan(no_need), include=FALSE}
colnames(death_confirm)[1] <- "date"
colnames(concern)[1] <- "date"
concern$date <- as.Date(concern$date, '%m/%d/%Y')
concern_long <- concern %>% pivot_longer(
  cols = dem:rep,
  names_to = "poli",
  values_to = "val"
)

concern_long_Daily_conf <- left_join(concern_long,confirm_total,by = c("date" = "Date"))

ggplot(concern_long,aes(x = date, y = val, color = poli))+
  geom_line()+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  labs(title = "Concerns over Time",
       subtitle  = "Positive = Concern, Negetive = Unconcern",
       x = "Date", y = "Concerns", color = "Politicians")+
  scale_colour_discrete(labels = c("Democrats", "Republicans"))
```
# Connection between concern and daily confirm?\
```{r, warning=FALSE, echo = FALSE}
coef <- 33725/108
ggplot(concern_long_Daily_conf,aes(x = date))+
  geom_line(aes(y = val, color = poli,size = ))+
  geom_line(aes(y = Daily_confirmed/coef), color="purple4")+
  scale_y_continuous(
    
    # Features of the first axis
    name = "Concerns",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~.*coef, name="Daily Confirmed")
  ) + 
  
  theme_ipsum() +

  theme(
    axis.title.y = element_text(color = "black", size=13),
    axis.title.y.right = element_text(color = "purple4", size=13),
    axis.line.y.right = element_line(color = "purple4"),
    axis.text.y.right = element_text(color = "purple4"))+
  geom_hline(yintercept=0, linetype="dashed", color = "black")+
  labs(title = "Concerns over Time",
       subtitle  = "Positive = Concern, Negetive = Unconcern",
       x = "Date", color = "Political opinion")+
  scale_colour_discrete(labels = c("Democrats", "Republicans"))+
  theme(legend.position="left")
```
\
\
```{r eval = FALSE,echo = FALSE}
# x <- confirm_dates_longer %>% mutate(
#   week = week(Date) + ifelse(year(Date)==2021,53,0),
#   month = month(Date),
#   year = year(Date)
# )
# x <- x %>% group_by(Country,week, month, year) %>% summarise(mean_confirm = mean(confirms))
# write.csv(confirm_dates_longer,"x.csv")
# # adding new column with function that sum infected in the last 14 days/2 weeks
```
## Spreading COVID among the countries through time\
* calculate daily active by excel equation\
* apply log function because the confirms rise exponential\
* norm the data BTW 0-1\
We want to show trends of Spreading\
```{r, echo = FALSE}
x <- read.csv("x.csv")
x <- rename(x,c("state"="Country"))
x <- x %>% mutate(corrent_conf = ifelse(ln(corrent_conf)<0,0,ln(abs(corrent_conf))))
x <- x %>% mutate(corrent_conf_norm = (corrent_conf-min(corrent_conf)/(max(corrent_conf)-min(corrent_conf))))
result <- plot_usmap(regions = "state", data = x, values = "corrent_conf_norm") +
  theme(panel.background = element_rect(colour = "black"))+
  scale_fill_distiller(palette = "Spectral") +
  labs(title = "week: {frame_time}",subtitle = "2 Waves, week 30 & week 48", caption = "week 30 ~ August 2020, week 48 ~ January 2021")+
    theme(legend.position = "right")+
  transition_time(week)
animate(result, fps = 4)
```





